An Implementation of the Active Contours without Edges model and the Logic Framework for Active contours on multi-channel images

### 2010-12-13 trunk/cs870/project/Report/Paper.tex
conc only missing

\usepackage{cvpr}
\usepackage{cite}
\usepackage{url}
\usepackage{graphicx}
\usepackage{clrscode}
\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{array}
\usepackage{color}
\usepackage{balance}

@@

\begin{abstract}
%\boldmath
The abstract goes here.
\end{abstract}

@@

\begin{equation}
\label{eqn:finters}
f_{\cup} = 1 = ((1 - z_{1}) .(1-  z_{2}))^{1/2}
\end{equation}

@@

\section{Difficulties and Discussion}
\label{sec:difficulties}

\section{Conclusion}
\label{sec:concl}
The conclusion goes here.

\bibliographystyle{IEEEtran}

\bibliography{references}

>>>>

\usepackage{cvpr}
\usepackage{cite}
\usepackage{url}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{array}
\usepackage{color}
\usepackage{balance}

@@

\begin{abstract}
In this project, we provide an implementation for the Chan-Vese model for active contours without edges, and the Sandberg-Chan logic framework for active
contours on multi-channel images. The Chan-Vese model is a special case of the more general Mumford-Shah functional for segmentation and level sets. It differs
from other active contour modesl in that it does not depend on the gradient of the contour ($\phi$), therefore it is more capable of detecting objects whose
boundaries may not be defined by a gradient (e.g. blurry images, noisy images). The Sandberg-Chan model builds on top of the Chan-Vese model by allowing the
contour to evolve simultaneously on multiple channels. It also offers a set of logic operations that can be applied to those channels (union, intersection,
complement, and their combinations). We evaluate our implementation based on several criteria, such as: independence of initial curve position, successful
detection of holes, performance with blurred and noisy images, and response to scaling parameters.
\end{abstract}

@@

\begin{equation}
\label{eqn:finters}
f_{\cap} = 1 = ((1 - z_{1}) .(1-  z_{2}))^{1/2}
\end{equation}

@@

\section{Difficulties and Discussion}
\label{sec:difficulties}
We have faced many challenges throughout the implementation of this project. Some of the challenges caused major delays in the implementation cycle and hindered
us from improving our code and adding more features to it.

\subsection{Varying $\mu$}
$\mu$ is a scaling factor that is used in both the Chan-Vese model and the Sandberg-Chan model to control the weight given to the length of the contour.
Therefore, varying the value of $\mu$ will affect the ability to segment objects in noisy images, especially when the noise is similar to Matlab's \textit{salt
and pepper} noise. However, varying $\mu$ in our implementation does not really have any impact on the length of the contour. Figure [\ref{fig:cvmu}] shows the
segmentation of an object in a noisy for different values for $\mu$. The resulting segmentation is the same for all values of $\mu$. It is worth noting that
this challenge kept us busy for over two days, after which we decided we should move on to other parts of the project.

\begin{figure*}[t]
\centering
\includegraphics[width=\textwidth]{cvmu.png}
\caption{$\mu$ does not have any effect on the length of the contour. \textit{Salt and pepper} noise was added to the original image for testing purposes. The
figure shows the segmentation for the same image for various values of $\mu$.}
\label{fig:cvmu}
\end{figure*}

\subsection{Computing the complement}
The Sandberg-Chan model supports the complement logic operator, so one can segment the logic object ($A \cap \neg B$) (i.e. segment what is in channel A and not
in channel B). The model defines the complement of the object in channel \textit{i} as follows:

\begin{equation}
\label{eqn:complement}
\begin{align*}
{z_{i}}^{in^{'}} = 1 - {z_{i}}^{in}, \\
{z_{i}}^{out^{'}} = 1 - {z_{i}}^{out}.
\end{align*}
\end{equation}

Although the definition is very simple, it does not work in practice. We implemented the complement logic operator as defined in Equation
[\ref{eqn:complement}], and the segmentation does not stop at the required result. It rather keeps evolving after reaching the required solution, then keeps
flipping between two segmentations: the correct segmentation, and the segmentation of $A \cap B$. Figure [\ref{fig:sccomp}] shows this scenrio. We tried to
debug the reasons for this problem for three days.

\begin{figure*}[t]
\centering
\includegraphics[width=\textwidth]{sccomp.png}
\caption{The complement operator causes the contour to go through an infinite loop between two segmentations: the correct segmentation, and the segmentation
of $A \cap B$.}
\label{fig:sccomp}
\end{figure*}

\subsection{Defining the initial contour ($\phi_0$)}
Both the Chan-Vese model and the Sandberg-Chan model defines the initial contour $\phi_0$ as a signed distance function where $\phi_0$ is positive inside the
contour $C$, zero on the boundary of $C$ (\partial C), and negative outside the contour $C$. Figure [\ref{fig:phinote}] illustrates those assumptions.

\begin{figure}[t]
\centering
\includegraphics[width=\textwidth]{phinote.png}
\caption{Assumptions about the way $\phi_0$ is defined.}
\label{fig:phinote}
\end{figure}

Since we model our initial contour as a circle, we have to modify the regular equation of the circle to match those assumption. Therefore, we initialize
$\phi_0$ as follows:

\begin{equation}
\label{eqn:phinote}
\phi_{0} (x, y) = radius - \sqrt{ (x - c_x)^2 + (y - c_y)^2 },
\end{equation}
where $c_x$ and $c_y$ are the x-coordinate and y-coordinate of the center of the circle respectively.

Although this definition of $\phi_0$ is compatible with most of the images we used to test our implementation, it did fail to correctly segment the objects in
some images. In fact, the results were the opposite of the used logic operator (i.e. the output is the intersection of the channels, while it should be the
union). After spending quite sometime to resolve this problem and checking what might be the source of the error, we tried using the regular equation of a
circle to initialize $\phi_0$ (Equation [\ref{eq:phinotenormal]}). We were very intrigued to notice that this modification resulted in the correct segmentation
of the input channels. Figure [\ref{fig:phidef}] shows the final segmentation using both definitions for $\phi_0$.

\begin{equation}
\label{eqn:phinotenormal}
\phi_{0} (x, y) = \sqrt{ (x - c_x)^2 + (y - c_y)^2 } - radius,
\end{equation}
where $c_x$ and $c_y$ are the x-coordinate and y-coordinate of the center of the circle respectively.

\begin{figure*}[t]
\centering
\includegraphics[width=\textwidth]{phidef.png}
\caption{}
\label{fig:phidef}
\end{figure*}

\subsection{Mistyped Equations}
As previously mentiond in Section [\ref{sec:sandberg-chan}], there is a typo in the equations that define $z_in$ and $z_out$ in the Sandberg-Chan model
(equation 3 in thier paper). In addition, they have another typo in the expansion of the the Eular-Lagrange equation for the PDE that models the problem. Both
typos consumed over a week of our time until we were able to find out what are the correct equations in both cases. We could have spent this time overcoming
other challenges that we faced while implementing both models.

\section{Conclusion}
\label{sec:concl}

\bibliographystyle{IEEEtran}
\bibliography{references}
